Are environmental and geographic variables effective surrogates for genetic variation in conservation planning?

### 2016-02-24 article/article.Rmd
+ update methods + replace % variation explained with 0 if less than 0

## Methods
### Study system
We used species distribution and genomic data collected by the IntraBioDiv project in the European Alps [@r451; see @r452 and @r478 for methodological details]. This dataset was particularly well suited for this study, because it provides spatially explicit genomic data for multiple species with vastly different demographic characteristics and evolutionary histories across a broad geographic area with large environmental gradients. As part of this project, data was collected for 27 alpine plant species using a 20' longitude by 21' latitude grid (approx. 21' 22.3 km $\times$ 25 km; Figure S1). They visited every second grid cell, and if a species was detected in a grid cell, samples were collected from three individuals. They genotyped samples using amplified fragment length polymorphisms [AFLP; @r453], and constructed matrices denoting the presence/absence of polymorphisms at loci for each species. Thus for each second grid cell, for each of the species present in the grid cell, the dataset contains information describing the genomic variation of individuals in that grid cell. We used these grid cells as planning units to develop prioritizations investigating the effectiveness of surrogates for the (AFLP) genetic variation. Note that we omitted grid cells that did not contain any individuals from our analysis. The majority of our analyses described hereafter were performed using the \texttt{R} software package [version `r paste0(version$major, '.', version$minor)`; @r192].

@@

### Outlier locus detection
To investigate the effectiveness of the surrogates for adaptive and neutral genetic variation, we first need to identify which of the sampled loci--if any--were under selection for each species. We used outlier locus detection methods [evaluated in @r454] to achieve this. The basic premise underpinning these analyses is that neutral loci are expected to exhibit a specific level of variation among populations ($F_{st}$), and loci that deviate from this expectation should be under selection. The advantage of these methods--in contrast with environmental association analyses [eg. r@475]--is that they do not use environmental data to identify loci under selection, which would have biased our analysis. In an earlier study, Bothwell _et al._ [-@r455] applied such outlier detection methods to one of the species in this dataset (_Gentiana nivalis_). Here, we adopted a similar methodology and applied it to each of the 27 species in the dataset.

First, we grouped individuals into genetic lineages for each species. The purpose of this step was to avoid misclassifying loci as adaptive. For each species, we used non-metric multi-dimensional scaling [NMDS; implemented in the $\texttt{vegan R}$ package] to transform the binary loci data into continuous variables (Table S2). We used Gower distances [@r464] to express the genomic differences between individuals [using the $\texttt{cluster R}$ package to accommodate sparsity; @r457]. We ran successive scaling analyses with increasing dimensionality until ordinations were obtained with an adequate stress value ($\leq$ `r nmds.params.LST[[MODE]]$max.stress`; `r nmds.params.LST[[MODE]]$trymax` random starts for all models). We then fit a suite of Gaussian mixture to the ordinated values to estimate the probability that individuals belonged to different populations [using the default priors and covariance structures implemented in the $\texttt{mclust} R$ package; @r477; @r476]. For each parametrization, we fit a set of models assuming varying numbers of total populations (\textit{G}= 1--`r max(mclust.params.LST[[MODE]]$G)`). We identified the most parsimonious models (lowest Bayesian Information Criterion $BIC$ statistic; Figures S4), and used these to group individuals into populations (Figure S5).

We identified loci under selection (Table S3). For each species, we ran $\texttt{Bayescan}$ [version 2.1; @r456] using the population memberships identified in the previous analysis (1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations). To avoid false-positive, we omitted loci where the global frequency of the minor allele was $\geq 0.05$ (following guidelines in the $\texttt{Bayescan}$ manual). We ran `r bayescan.params.LST[[MODE]]$reps` replicates per species using a suitable false discovery rate to ensure convergence [FDR $\leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466].

After classifying loci as adaptive or neutral, we mapped the main gradients of the adaptive (if detected) and neutral genetic variation for each species (Figures S6--S32). For each species, we partitioned adaptive and neutral loci into separate datasets, and reran the nonmetric multidimensional scaling analysis (NMDS) to derive continuous variables that described the main gradients of adaptive and neutral genetic variation separately (Table S2). Since each grid cell had up to three samples per species, we used the average of the ordinated values associated with the individuals in each grid cell to summarize the typical genetic characteristics of individuals in the cells.

>>>>

## Methods
### Study system
We used species distribution and genomic data collected by the IntraBioDiv project in the European Alps [@r451; see @r452 and @r478 for methodological details]. This data set was particularly well suited for this study, because it provides spatially explicit genomic data for multiple species with vastly different demographic characteristics and evolutionary histories across a broad geographic area with large environmental gradients. As part of this project, data was collected for 27 alpine plant species using a 20' longitude by 21' latitude grid (approx. 21' 22.3 km $\times$ 25 km; Figure S1). They visited every second grid cell, and if a species was detected in a grid cell, samples were collected from three individuals. They genotyped samples using amplified fragment length polymorphisms [AFLP; @r453], and constructed matrices denoting the presence/absence of polymorphisms at loci for each species. Thus for each second grid cell, for each of the species present in the grid cell, the data set contains information describing the genomic variation of individuals in that grid cell. We used these grid cells as planning units to develop prioritizations investigating the effectiveness of surrogates for the (AFLP) genetic variation. Note that we omitted grid cells that did not contain any individuals from our analysis. The majority of our analyses described hereafter were performed using the \texttt{R} software package [version `r paste0(version$major, '.', version$minor)`; @r192].

@@

### Outlier locus detection
To investigate the effectiveness of the surrogates for adaptive and neutral genetic variation, we first need to identify which of the sampled loci--if any--were under selection for each species. We used outlier locus detection methods [evaluated in @r454] to achieve this. The basic premise underpinning these analyses is that neutral loci are expected to exhibit a specific level of variation among populations ($F_{st}$), and loci that deviate from this expectation should be under selection. The advantage of these methods--in contrast with environmental association analyses [eg. r@475]--is that they do not use environmental data to identify loci under selection, which would have biased our analysis. In an earlier study, Bothwell _et al._ [-@r455] applied such outlier detection methods to one of the species in this data set (_Gentiana nivalis_). Here, we adopted a similar methodology and applied it to each of the 27 species in the data set.

First, we grouped individuals into genetic lineages for each species (using a methodology similar to that implemented in \texttt{prabclust R} package; @r479). The purpose of this step was to avoid misclassifying loci as adaptive. For each species, we ordinated the genomic data and fit models to the ordinations to group individuals into populations. Specifically, we used non-metric multi-dimensional scaling [NMDS; implemented in the $\texttt{vegan R}$ package] to transform the binary loci data into continuous variables (Table S2). We used Gower distances [@r464] to express the genomic differences between individuals [using the $\texttt{cluster R}$ package to accommodate sparsity; @r457]. To ensure that the ordinations described an adequate amount of the genomic variation, we ran successive scaling analyses with increasing dimensionality to obtained ordinations until this was achieved (stress value $\leq$ `r nmds.params.LST[[MODE]]$max.stress`; `r nmds.params.LST[[MODE]]$trymax` random starts for each analysis). We then fit a suite of Gaussian mixture to the ordinated values to estimate the probability that individuals belonged to different populations [using the default priors and covariance structures implemented in the $\texttt{mclust} R$ package; @r477; @r476]. For each parametrization, we fit a set of models assuming varying numbers of total populations (\textit{G}= 1--`r max(mclust.params.LST[[MODE]]$G)`). We identified the most parsimonious models (lowest Bayesian Information Criterion $BIC$ statistic; Figures S4), and used these to group individuals into populations (Figure S5).

We identified loci under selection (Table S3). Since outlier locus detection methods compare allele frequencies between multiple populations, we were unable to investigate species that only contained a single population. For these species ($n = `r sum(sapply(spp.Mclust.LST, function(x) return(ncol(x$mclust$z)==1)))`$), we assumed that all genetic variation was neutral. For species with more than two populations, we ran $\texttt{Bayescan}$ [version 2.1; @r456; 1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations]. To avoid false-positives, we omitted loci where the global frequency of the minor allele was $\geq 0.05$ (following guidelines in the $\texttt{Bayescan}$ manual). We ran `r bayescan.params.LST[[MODE]]$reps` replicates per species using a suitable false discovery rate to ensure convergence [FDR $\leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466].

After classifying loci as adaptive or neutral, we mapped the main gradients of the adaptive (if detected) and neutral genetic variation for each species (Figures S6--S32). For each species, we partitioned adaptive and neutral loci into separate data sets, and reran the nonmetric multidimensional scaling analysis (NMDS) to derive continuous variables that described the main gradients of adaptive and neutral genetic variation separately (Table S2). Since each grid cell had up to three samples per species, we used the average of the ordinated values associated with the individuals in each grid cell to summarize the typical genetic characteristics of individuals in the cells.
